---
title: "OptGoal"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{OptGoal}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(OptGoal)
```
## Introduction

Water reservoir management is one of the important topics in operations research due to its practical necessity. There are many techniques available in the literature of operations research that help in the decision-making of water reservoir management. The main objective of this study is to create a R package by implementing goal programming to optimize water release. 

While there are existing packages that address water release optimization, focus solely on linear programming 
methods without incorporating goal programming. This package fills this gap by providing tools for goal programming optimization and reliability assessment in water reservoir management. Furthermore, the package includes visualization functionalities to effectively display 
input data and output results, for better interpretation and decision-making. This package is based on a case study. 

## Model

#### Inputs

1. Monthly average inflow to the reservoir ($I_t$)
2. Monthly water demand ($D_t$)
3. Monthly minimum storage of the reservoir ($S_{min}$)
4. Monthly maximum storage of the reservoir ($S_{max}$)

#### Deviation variables 

1. Positive deviation ($d_i^+$)
2. Negative deviation ($d_i^-$)

#### Objective function

Consider $R_t$ $\leq$ $D_t$ for $t$ = 1,...,12.

So, the negative deviations must be minimized.
\[Minimize: Z = \sum_{i=1}^n d_i^- \]

#### Constraints

1. System Constraints
   i. Continuity Equation
  $$S_{t+1} = S_t + I_t - R_t$$ 
   ii. Storage limits
   $$S_{min} \leq S_t \leq  S_{max}$$
   
2. Goal Constraints
$$R_t - d_i^+ + d_i^- = D_t$$
3. Non - negativity constraints
$$R_t \geq 0$$
$$S_t \geq 0 $$
$$d_i^+ \geq 0$$
$$d_i^- \geq 0 $$

## Data (1996 - 2020)

Data collected from 1996 to 2020. In this model, we use the average of the data. Data source: Mahaweli Authority of Sri Lanka.

| Month    | Inflow ($m^3$) | Demand ($m^3$)   | Minimum_Storage ($m^3$) | Maximum_Storage ($m^3$) |
|---------|-------|--------------|----------------- |----------------|
| January   | 7569  | 11586.72     | 2824.55 | 26864.25 |
| February     | 6383  | 4057.81 | 2824.55 | 26864.25|
| March | 8472  | 3586.72      | 2824.55| 26864.25 | 
| April | 13084  | 10057.81      | 2824.55| 26864.25 | 
| May | 9689  | 3456.72      | 2824.55| 26864.25 | 
| June | 4301  | 1057.81      | 2824.55| 26864.25 | 
| July | 399  | 6586.72      | 2824.55| 26864.25 | 
| August | 2896  | 2057.81      | 2824.55| 26864.25 | 
| September | 2723  | 1096.72      | 2824.55| 26864.25 |
| October | 7239  | 1057.81      | 2824.55| 26864.25 | 
| November | 14849  | 1186.72      | 2824.55| 26864.25 | 
| December | 13304  | 1808.81      | 2824.55| 26864.25 |
|          |         |            |         |          |

## optimize_release function

This R function was created by including relevant inputs and defining the constraints that describe the model. This function provides the following outputs:

1. Monthly release ($R_t$)
2. Monthly storage ($S_t$)
3. Positive deviation ($d_i^+$)
4. Negative deviation ($d_i^-$) 
5. Objective value

Required library: lpSolve.

#### R code (optimize_release.R)

```{r}
library(lpSolve)
optimize_release <- function(inflow, demand, constraints) {
  T <- length(inflow)  # Number of time periods

  # Input validation
  if (length(demand) != T) {
    stop("Inflow and demand vectors must be of the same length.")
  }
  if (length(constraints$S_min) != T || length(constraints$S_max) != T) {
    stop("S_min and S_max must be vectors of length equal to the number of time periods.")
  }
  if (any(constraints$S_min > constraints$S_max)) {
    stop("Each element of S_min must be less than or equal to the corresponding element in S_max.")
  }


  # Decision variables for each time period:
  # R_t (release), S_t (storage), d_t^+, d_t^-
  num_vars <- T * 4  # Total number of variables

  # Objective function coefficients
  obj_coeffs <- c(
    rep(0, T * 2),                 # Zero coefficients for R_t and S_t
    rep(1, T),                     # Weights for d_t^+
    rep(1, T)                      # Weights for d_t^-
  )

  # Initialize constraints matrices
  constraints_matrix <- matrix(0, nrow = 0, ncol = num_vars)
  constraints_direction <- character(0)
  constraints_rhs <- numeric(0)

  # Constraint indices
  idx_R <- 1:T
  idx_S <- (T + 1):(2 * T)
  idx_d_plus <- (2 * T + 1):(3 * T)
  idx_d_minus <- (3 * T + 1):(4 * T)

  # 1. Storage Continuity Equations
  for (t in 1:T) {
    row <- rep(0, num_vars)
    if (t == 1) {
      # For t = 1: S_0 + I_1 - R_1 - S_1 = 0
      row[idx_S[t]] <- -1       # -S_1
      row[idx_R[t]] <- -1       # -R_1
      rhs <- -constraints$initial_storage - inflow[t]
    } else {
      # For t > 1: S_{t-1} + I_t - R_t - S_t = 0
      row[idx_S[t - 1]] <- 1    # S_{t-1}
      row[idx_S[t]] <- -1       # -S_t
      row[idx_R[t]] <- -1       # -R_t
      rhs <- -inflow[t]
    }
    constraints_matrix <- rbind(constraints_matrix, row)
    constraints_direction <- c(constraints_direction, "=")
    constraints_rhs <- c(constraints_rhs, rhs)
  }

  # 2. Goal Constraints
  for (t in 1:T) {
    row <- rep(0, num_vars)
    row[idx_R[t]] <- 1          # R_t
    row[idx_d_plus[t]] <- -1     # +d_t^+
    row[idx_d_minus[t]] <- 1   # -d_t^-
    constraints_matrix <- rbind(constraints_matrix, row)
    constraints_direction <- c(constraints_direction, "=")
    constraints_rhs <- c(constraints_rhs, demand[t])
  }

  # 3. Storage Limits (Lower Bounds)
  for (t in 1:T) {
    row <- rep(0, num_vars)
    row[idx_S[t]] <- 1  # S_t
    constraints_matrix <- rbind(constraints_matrix, row)
    constraints_direction <- c(constraints_direction, ">=")
    constraints_rhs <- c(constraints_rhs, constraints$S_min[t])
  }

  # 4. Storage Limits (Upper Bounds)
  for (t in 1:T) {
    row <- rep(0, num_vars)
    row[idx_S[t]] <- 1  # S_t
    constraints_matrix <- rbind(constraints_matrix, row)
    constraints_direction <- c(constraints_direction, "<=")
    constraints_rhs <- c(constraints_rhs, constraints$S_max[t])
  }

  # 5. Release Bounds (Non-negativity)
  for (t in 1:T) {
    row <- rep(0, num_vars)
    row[idx_R[t]] <- 1  # R_t
    constraints_matrix <- rbind(constraints_matrix, row)
    constraints_direction <- c(constraints_direction, ">=")
    constraints_rhs <- c(constraints_rhs, 0)
  }

  # 6. Deviations Non-negativity
  # d_t^+ >= 0
  for (t in 1:T) {
    row <- rep(0, num_vars)
    row[idx_d_plus[t]] <- 1  # d_t^+
    constraints_matrix <- rbind(constraints_matrix, row)
    constraints_direction <- c(constraints_direction, ">=")
    constraints_rhs <- c(constraints_rhs, 0)
  }
  # d_t^- >= 0
  for (t in 1:T) {
    row <- rep(0, num_vars)
    row[idx_d_minus[t]] <- 1  # d_t^-
    constraints_matrix <- rbind(constraints_matrix, row)
    constraints_direction <- c(constraints_direction, ">=")
    constraints_rhs <- c(constraints_rhs, 0)
  }

  # Solve the optimization problem
  result <- lp(
    direction = "min",
    objective.in = obj_coeffs,
    const.mat = constraints_matrix,
    const.dir = constraints_direction,
    const.rhs = constraints_rhs
  )

  if (result$status != 0) {
    print(paste("LP Status Code:", result$status))
    stop("Optimization did not find a feasible solution.")
  }

  # Extract the optimized release and storage
  optimized_release <- result$solution[idx_R]
  optimized_storage <- result$solution[idx_S]
  deviations_positive <- result$solution[idx_d_plus]
  deviations_negative <- result$solution[idx_d_minus]

  # Return the results
  return(list(
    release = optimized_release,
    storage = optimized_storage,
    deviations_positive = deviations_positive,
    deviations_negative = deviations_negative,
    objective_value = result$objval
    )
  )
}
#Run the example
inflow <- c(7569, 6383, 8472, 13084, 9689, 4301, 399, 2896, 2723, 7239,14849, 13304)
demand <- c(11586.72, 4057.81, 3586.72, 10057.81,
3456.72, 1057.81, 6586.72, 2057.81, 1096.72, 1057.81, 1186.72, 1808.81)
S_min_vector <- rep(2824.55, time=12)
S_max_vector <- rep(26864.25, time=12)
constraints <- list(initial_storage = 500, S_min = S_min_vector, S_max = S_max_vector)
result <- optimize_release(inflow, demand, constraints)
result
```

Here, we manually enter the data. Additionally, data can be defined as follows, called, and executed.

```{r}
reservoir_data <- data.frame(
  Month = month.abb,
  Inflow = c(7569, 6383, 8472, 13084, 9689, 4301, 399, 2896, 2723, 7239, 14849, 13304),
  Demand = c(11586.72, 4057.81, 3586.72, 10057.81, 3456.72, 1057.81, 6586.72, 2057.81, 1096.72, 1057.81, 1186.72, 1808.81),
  S_min = rep(2824.55, 12),
  S_max = rep(26864.25, 12)
)
reservoir_data
```
## assess_reliability function

This R function was created by including demand and release, which were obtained by previous code. This function provides the following outputs:

1. Time reliability
2. Volume reliability

#### R code (assess_reliability.R)

```{r}
assess_reliability <- function(release, demand) {
  if (length(release) != length(demand)) {
    stop("Release and demand vectors must be of the same length.")
  }
  T <- length(release)

  # Time-Based Reliability
  time_reliability <- sum(release >= demand) / T

  # Volume-Based Reliability
  volume_reliability <- sum(release) / sum(demand)

  # Compile results
  reliability_results <- data.frame(time_reliability = time_reliability, volume_reliability = volume_reliability)

  return(reliability_results)
}
#Run the example
inflow <- c(7569, 6383, 8472, 13084, 9689, 4301, 399, 2896, 2723, 7239,14849, 13304)
demand <- c(11586.72, 4057.81, 3586.72, 10057.81,
3456.72, 1057.81, 6586.72, 2057.81, 1096.72, 1057.81, 1186.72, 1808.81)
S_min_vector <- rep(2824.55, time=12)
S_max_vector <- rep(26864.25, time=12)
constraints <- list(initial_storage = 500, S_min = S_min_vector, S_max = S_max_vector)
result <- optimize_release(inflow, demand, constraints)
release <- result$release
reliability_results <- assess_reliability(release, demand)
reliability_results
```

## plot_release function

This function provides a graphical representation of the release in order to months. x-axis denotes the months, and y-axis denotes release in cubic meters. Required library: ggplot2.

#### R code (plot_release.R)

```{r plot_release, fig.width=5, fig.height=3, dpi=120}
library(ggplot2)
plot_release <- function(release, months = NULL) {
  if (!is.numeric(release)) {
    stop("release must be a numeric vector.")
  }

  time <- seq_along(release)

  # Assign default month names if appropriate
  if (is.null(months) && length(release) == 12) {
    months <- month.abb
  }

  # Use months if provided
  if (!is.null(months)) {
    if (length(months) != length(release)) {
      stop("Length of 'months' must match length of 'release'.")
    }
    time_labels <- months
  } else {
    time_labels <- time
  }

  data <- data.frame(
    Time = time,
    Release = release
  )

  plot <- ggplot2::ggplot(data, ggplot2::aes(x = .data$Time, y = .data$Release)) +
    ggplot2::geom_line(color = "red", linewidth = 1) +
    ggplot2::geom_point(color = "red") +
    ggplot2::scale_x_continuous(breaks = time, labels = time_labels) +
    ggplot2::theme_minimal() +
    ggplot2::labs(
      title = "Release Over Time",
      x = "Month",
      y = "Release Volume (Cubic meters)"
    )

  return(plot)
}
#Run the example
inflow <- c(7569, 6383, 8472, 13084, 9689, 4301, 399, 2896, 2723, 7239,14849, 13304)
demand <- c(11586.72, 4057.81, 3586.72, 10057.81,
3456.72, 1057.81, 6586.72, 2057.81, 1096.72, 1057.81, 1186.72, 1808.81)
S_min_vector <- rep(2824.55, time=12)
S_max_vector <- rep(26864.25, time=12)
constraints <- list(initial_storage = 500, S_min = S_min_vector, S_max = S_max_vector)
result <- optimize_release(inflow, demand, constraints)
release <- result$release
plot_release(release)
```

## plot_release_vs_demand function

This function provides a graphical representation of the release and demand in order to months. x-axis denotes the months, and y-axis denotes volume in cubic meters. Required library: ggplot2.

#### R code (plot_release_vs_demand.R)

```{r plot_release_vs_demand, fig.width=5, fig.height=3, dpi=120}
library(ggplot2)
plot_release_vs_demand <- function(release, demand, months = NULL) {
  if (!is.numeric(release)) {
    stop("release must be a numeric vector.")
  }
  if (!is.numeric(demand)) {
    stop("demand must be a numeric vector.")
  }
  if (length(release) != length(demand)) {
    stop("release and demand must be of the same length.")
  }

  time <- seq_along(release)

  # Assign default month names if appropriate
  if (is.null(months) && length(release) == 12) {
    months <- month.abb  # Built-in vector of abbreviated month names
  }

  # Use months if provided
  if (!is.null(months)) {
    if (length(months) != length(release)) {
      stop("Length of 'months' must match length of 'release' and 'demand'.")
    }
    time_labels <- months
  } else {
    time_labels <- time
  }

  data <- data.frame(
    Time = time,
    Release = release,
    Demand = demand
  )

  # Melt data for plotting
  data_melted <- reshape2::melt(data, id.vars = "Time")

  # Create the plot
  plot <- ggplot2::ggplot(data_melted, ggplot2::aes(x = .data$Time, y = .data$value, color = .data$variable)) +
    ggplot2::geom_line(linewidth = 1) +
    ggplot2::geom_point() +
    ggplot2::scale_x_continuous(breaks = time, labels = time_labels) +
    ggplot2::theme_minimal() +
    ggplot2::labs(
      title = "Release vs. Demand Over Time",
      x = "Month",
      y = "Volume (Cubic meters)",
      color = "Legend"
    )

  return(plot)
}
#Run the example
inflow <- c(7569, 6383, 8472, 13084, 9689, 4301, 399, 2896, 2723, 7239,14849, 13304)
demand <- c(11586.72, 4057.81, 3586.72, 10057.81,
3456.72, 1057.81, 6586.72, 2057.81, 1096.72, 1057.81, 1186.72, 1808.81)
S_min_vector <- rep(2824.55, time=12)
S_max_vector <- rep(26864.25, time=12)
constraints <- list(initial_storage = 500, S_min = S_min_vector, S_max = S_max_vector)
result <- optimize_release(inflow, demand, constraints)
release <- result$release
plot_release_vs_demand(release, demand)
```

## Advantages

Finding a workable solution in a linear programming model is impossible if the data does not meet the restrictions. However, goal programming allows us to identify the ways in which our solution differs from the workable option. Therefore, we may effectively tackle our problem by utilizing this model in conjunction with the R program.

## References

[1] http://www.pgis.pdn.ac.lk/rescon2021/abstracts/ICTMS/44.pdf

[2] https://citeseerx.ist.psu.edu/document?repid=rep1&type=pdf&doi=ab4bd579b616522619686295eaee77bfc17745a8

[3] https://www.researchgate.net/publication/28600555_Optimal_Reservoir_Operation_Using_Multi-Objective_Evolutionary_Algorithm
