---
title: "OptGoal"
author: "Vijini Lakmini, Rathnayake Mudiyanselage"
date: "`r Sys.Date()`"
output: pdf_document
vignette: >
  %\VignetteIndexEntry{An Introduction to OptGoal}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
   collapse = TRUE)
```

```{r setup, echo = FALSE}
library(OptGoal)
```
Water reservoir management is one of the important topics in operations research due to its practical necessity. There are many techniques available in the literature of operations research that help in decision making of water reservoir management. The main objective of this study is to create a R package by implementing goal programming to optimize water release. 

While there are existing packages that address water release optimization, focus solely on linear programming 
methods without incorporating goal programming. This package fills this gap by providing tools for goal programming optimization and reliability assessment in water reservoir management. Furthermore, the package includes visualization functionalities to effectively display 
input data and output results, for better interpretation and decision-making. 

```{r}

```

